Convolutional Neural Networks on Graphs\\ with Fast Localized Spectral Filtering

### 2016-10-29 nips2016/nips2016.tex
nips v5: large edition by Michaël (2016-06-17)

\usepackage{nips2016}
%\usepackage[final]{nips_2016}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amssymb}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{color}

@@

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  Michaël Defferrard \\ %\thanks{further info} \\
  %Department of Electrical Engineering \\
  %École Polytechnique Fédérale de Lausanne (EPFL) \\
  EPFL \\
  Lausanne, Switzerland \\
  \texttt{michael.defferrard@epfl.ch} \\
  \And % \AND
  Xavier Bresson \\
  EPFL \\
  Lausanne, Switzerland \\
  \texttt{xavier.bresson@epfl.ch} \\
  \And
  Pierre Vandergheynst \\
  EPFL \\
  Lausanne, Switzerland \\
  \texttt{pierre.vandergheynst@epfl.ch} \\
}

\usepackage[acronym]{glossaries}
\newacronym{SVD}{SVD}{Singular Value Decomposition}
\newacronym{SGD}{SGD}{Stochastic Gradient Descent}
\newacronym{PSD}{PSD}{positive semidefinite}

\DeclareMathOperator*{\diag}{diag}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\spn}{span}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\W}{\mathcal{W}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Xh}{\hat{X}}
\newcommand{\Yh}{\hat{Y}}
\newcommand{\st}{\ \text{s.t.} \,}
\newcommand{\norm}[1]{\left\| #1 \right\|}

\begin{document}
% \nipsfinalcopy is no longer used

@@

\begin{abstract}
Convolutional neural networks have greatly improved state-of-the-art performances in computer vision and speech analysis tasks, due to its high ability to extract multiple levels of representations of data. In this work, we are interested in generalizing convolutional neural networks from low-dimensional regular grids, where image, video and speech are represented, to high-dimensional irregular domains, such as social networks, biological graphs like gene regulatory and brain connectivity networks, telecommunication networks, or words' embedding. We present a formulation of convolutional neural networks on graphs in the context of the emerging field of signal processing on graphs, which provides the necessary mathematical background and efficient numerical schemes to design fast localized convolutional filters on graphs. Numerical experiments on MNIST and 20NEWS demonstrate the ability of the system to learn local stationary features on graphs, as long as the graph is well-constructed.
\end{abstract}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=13.5cm]{images/illustrationCNNgraphs.eps}
%\includegraphics[width=13.5cm]{images/illustrationCNNgraphs.png}
\caption{Architecture of the proposed CNNs on graphs. Notation: $l$ is the coarsening level, $x^l$ are the downsampled signals at layer $l$, $G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the coarsening exponent, $n_c$ is the number of classes, $y$ is the output signal, and $\theta^l$ is the number of parameters to learn at $l$.}
\label{fig:}
\end{figure}

\section{Introduction}
\vspace{-0.2cm}
Convolutional neural networks (CNNs) \cite{pro:LeCunBottouBengioHaffner98MNIST} offer an efficient architecture to extract highly meaningful statistical patterns in large-scale and high-dimensional datasets. The key success of CNNs is its advanced ability to learn local stationary structures and compose them to form multi-scale hierarchical patterns. Precisely, CNNs extract the local stationarity property of the input data or signals by revealing local features that are shared across the data domain. These similar features are identified with localized convolutional filters or kernels, which are learned from the data. Convolutional filters are shift- or translation-invariant filters, meaning they are able to recognize identical features independently of their spatial locations. Localized kernels or compactly supported filters refer to filters that extract local features independently of the input data size, with a support size that can be much smaller than the input size. \\
The compositional local stationarity property of CNNs can be efficiently implemented on regular grids, like image and sound domains, as convolutional and downsampling operators are mathematically well-defined on such discretized Euclidean spaces. This has led to the breakthrough of CNNs in image, video, and sound recognition tasks \cite{art:LeCunBengioHinton15DL} as these data lie on low-dimensional regular lattices. But not all data lie on regular grids. User data on social networks, gene data on biological regulatory networks, log data on telecommunication networks, or text documents on words' embedding are important examples of data lying on irregular or non-Euclidean domains. Unfortunately, CNNs cannot be directly applied to such complex domains.\\
A generalization of CNNs to irregular grids is not direct as the basic convolutional and downsampling operators, necessary to extract compositional local stationarity features, are not mathematically clear on these domains. This makes this extension mathematically challenging, both in theory and in implementation. The goal of this work is to leverage the recent works of \cite{pro:GregorLeCun10LRF,pro:CoatesNg11LRF,art:BrunaZarembaSzlamLeCun13DLgraphs,art:HenaffBrunaLeCun15DLgraphs,art:HammondVandergheynstGribonval11GraphWav,art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG,art:ShumanRicaudVandergheynst16Local} to introduce an efficient implementation of CNNs on irregular domains represented here as graphs. Graphs are universal representations of heterogeneous pairwise relationships between possibly high-dimensional data. Graphs can encode complex geometric structures, and can be studied with strong mathematical tools such as graph spectral theory, a blend between graph theory and harmonic analysis.\\
The major bottleneck of generalizing CNNs to graphs is the design of fast and localized filters on graphs, that are the core elements to extract local stationarity structures. Our major goal is to overcome this barrier. We will make use of recent tools developed in the context of signal processing on graphs (SPG) \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} to precisely control the spatial support of graph filters using the so-called polynomial kernel localization \cite{art:ShumanRicaudVandergheynst16Local}, and to perform an efficient backward propagation technique using the Chebyshev technique \cite{art:HenaffBrunaLeCun15DLgraphs}. Eventually, a generic and efficient implementation of CNNs on graphs may pave the way to the development of a potentially large class of deep learning techniques for graph-based data.

\section{Related Works}
\vspace{-0.2cm}
{\bf CNNs on non-Euclidean domains.} Extending the success of CNNs to non-Euclidean domains is essential to boost the power analysis of data lying on complex networks like biological, social or telecommunication networks. This is a recent line of works, with a preliminary approach coming from the local reception fields in \cite{pro:GregorLeCun10LRF,pro:CoatesNg11LRF}, and applied successfully to object recognition in computer vision. The main idea is to unsupervisedly learn groups of similar features. This is highly useful to automatically choose a limited number of connections between two successive layers. This approach is however not able to extract similar groups over the data domain, as it is not clear how to define convolutional filters on these groups.  \\
In \cite{art:BrunaZarembaSzlamLeCun13DLgraphs,art:HenaffBrunaLeCun15DLgraphs}, the authors proposed a generalization of CNNs on graphs using graph counterparts of grid convolution and downsampling. They were able to learn shift-invariant or convolutional filters on graphs using the convolution theorem \cite{book:Mallat99wavelets} that defines convolutional operators as linear operators that diagonalize in the Fourier basis (represented by the eigenvectors of the Laplacian operator). They also designed a strategy to learn the graph structure from the data, and apply their model on image recognition, text categorization and bioinformatics. This promising approach is however limited in implementation as forward and backward evaluations have complexity $O(n^2)$ due to the matrix multiplication operations that operate the graph Fourier transforms. Besides, there is no strong control over the support size of the convolutional kernels, which is essential to design localized filters to extract local stationary features, independent of the input size. \\
The authors in \cite{pro:MasciBoscainiBronsteinVandergheynst15GeoDL,art:MasciBoscainiBronsteinVandergheynst15ShapeNet} introduced ShapeNet, a generalization of CNNs to 3D-meshes, a class of smooth low-dimensional non-Euclidean spaces. They used geodesic polar coordinates to define mesh patches and formulated a deep learning architecture. They obtained state-of-the-art results for shape recognition tasks.\\
Finally, the papers \cite{pro:ChenChengMallat14deepHaar,pro:RustamovGuibas14deepHaar} investigated the construction Haar wavelet transforms on graphs using a deep hierarchical architecture. They applied these deep wavelets on graphs to object recognition on sphere, and to sparse reconstruction of faces. \\
{\bf Signal Processing on Graphs.} SPG is an emerging field that aims at bridging the gap between signal processing techniques like wavelet analysis and graph theory such as spectral graph analysis \cite{art:BelkinNiyogi05LaplaBeltrami,art:VonLuxburg07Tutorial}. The main goal is to generalize fundamental analysis operations for signals from regular grids to irregular structures embodying by graphs. We refer the reader to \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} for an introduction of the field. Standard operations on grids such as convolution, translation, filtering, dilatation, modulation, downsampling do not extend directly to graphs and thus require new mathematical definitions while keeping the original intuitive concepts. In this context, the authors of \cite{art:HammondVandergheynstGribonval11GraphWav,art:CoifmanLafon06DifMap,pro:GavishNadlerCoifman10GraphHaar} revisited the construction of wavelet operators on graphs. In \cite{art:ShumanFarajiVandergheynst16PyramTrans,art:RamEladCohen11TreeWavelets}, the authors designed a technique to perform mutli-scale pyramid transforms on graphs. The works of \cite{pro:TsitsveroBarbarossa15Uncert,pro:PasdeloupAlamiGriponRabbat15Uncert,art:PerraudinRicaudShumanVandergheynst16Uncert} redefined uncertainty principles on graphs, and showed that intuitive concepts may be lost, but can also produce enhanced localization principles for signals on graphs. In \cite{pro:HammondRaoaroorJacquesVandergheynst10LassoGraWav}, it was shown how to carry out lasso-based signal regularization on graphs, and studied the intertwined relationships between smoothness and sparsity on graphs. In \cite{pro:TremblayPuyGribonvalVandergheynst16CompSpecClus}, the authors investigated compressed sensing recovery conditions for graph spectral signal analysis. \\
Beyond this work, our long-term goal is to make use of SPG tools as much as possible to formulate mathematically well-designed techniques of CNNs on graphs. This paper introduces the mathematical and computational foundations of such models, while future developments may consider enhanced graph filters like graph wavelet operators, spectral graph coarsening techniques, and improved localization properties.

\section{Contributions}
\vspace{-0.2cm}
{\textbf {\textit {(1)}}} Learning deep fully connected networks is almost computationally untractable for large data domain. Fortunately, for data satisfying stationarity and hierarchical compositionality structure, CNNs have proven to be remarkably powerful to identify highly informative features using low learning complexity. For example, learning a single layer with $n$ input coordinates and $m<n$ outputs require $O(n^2)$ complexity to learn fully connected networks. It also requires the complexity $O(n)$ to learn any arbitrary filter, and $O(S)$ for localized filters, where $S$ is the size of the filter support, independent of the input size $n$. One of our main contributions is to stay with the same low complexity for learning graph convolutional filters. We will show it is actually possible to keep exactly the same complexity as standard CNNs, i.e. $O(S)$ for localized graph filters.\\
{\textbf {\textit {(2)}}} The backpropagation learning algorithm requires to run convolutional operations, which can be efficiently implemented by FFT on grids with complexity $O(n\log n)$. Convolution on graphs can also be implemented by matrix multiplications representing the forward and backward graph Fourier transforms, but it is expensive as the complexity is $O(n^2)$. As noticed in \cite{art:HenaffBrunaLeCun15DLgraphs}, this complexity is a strong bottleneck to deal with large-scale problems. We will greatly improve this complexity by reducing to $O(E.S)$, where $E$ is the number of edges in the given graph, using the Chebyshev technique from SPG \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}. Besides, as most real-world graphs are highly sparse, we have $E\ll n^2$, and therefore a low complexity for the matrix multiplication operations. Eventually, the Chebyshev computations can be naturally distributed on GPUs, as we did in our experiments.\\
{\textbf {\textit {(3)}}} We introduce an efficient pooling strategy for CNNs on graphs, that also makes use of distributed computing architecture like GPUs. The pooling approach consists on two stages. First, a graph coarsening procedure based on METIS \cite{art:DhillonGuanKulis07Graclus,art:KarypisKumar98Metis} is successively applied to the original graph to produce coarser versions of the graph. Second, at each coarsening level, we sequentially order all nodes such that two successive indexed nodes are the parent nodes of the child node in the next coarser level. This node indexation structure in memory satisfies parallel architectures. \\
{\textbf {\textit {(4)}}} We regard this proposed work as the funding paper that bridges the gap between two powerful fields, namely deep learning and signal processing on graphs. In the future, we expect to further benefit from tools in SPG to enhance the formulation of CNNs on graphs. \\
{\textbf {\textit {(5)}}} We revisit the Euclidean case in the numerical experiments with the standard MNIST image classification problem. The model is able to learn localized convolutional filters and achieve 98.65\% of success rate. More interestingly, we apply our model to the text categorization problem on 20NEWS. We also demonstrate that our techniques is able to extract local stationary features, as it improves by 2.5\% the classification accuracy when no convolutional are present. We observe that the quality of the graph construction is strongly linked to the quality of the convolutional layers. Finally, we demonstrate the computational performance of our model over \cite{art:HenaffBrunaLeCun15DLgraphs}.

\section{Learning Localized Graph Filters}
\vspace{-0.2cm}
In this section, we explore how to design efficient localized convolutional filters on graphs.

\subsection{Signal Processing on Graphs and Localized Filters}
\vspace{-0.2cm}
{\bf Graph Fourier Transform.} We are interested in processing signals defined on undirected and connected graphs $G=(V,E,W)$, where $V$ is the set of vertices with $|V|=n$, $E$ is the set of edges and $W$ is the weighted adjacency matrix encoding the connection weights between two vertices. A signal $x:V\rightarrow\mathbb{R}$ defined on the nodes of the graph may be regarded as a vector $x\in\mathbb{R}^n$ where $x_i$ is the value of $x$ at the $i^{th}$ node. An essential operator in spectral graph analysis is the graph Laplacian \cite{book:Chung97Spectral}, which unnormalized definition is $L=D-W\in\mathbb{R}^{n\times n}$ with $D\in\mathbb{R}^{n\times n}$ is the diagonal degree matrix where $D_{ii}=\sum_jW_{ij}$, and its normalized definition is $L=I_n-D^{-1/2}WD^{-1/2}$ where $I_n$ is the identity matrix. As $L$ is a real symmetric positive semidefinite matrix, it has a complete set of orthonormal eigenvectors $\{u_l\}_{l=0}^{n-1}\in\mathbb{R}^n$, known as the graph Fourier modes, and their associated real nonnegative eigenvalues $\{\lambda_l\}_{l=0}^{n-1}$, which are the frequencies of the graph system. Equivalently, linear operator $L$ diagonalizes the Fourier basis matrix $U=[u_0,...,u_{n-1}]\in\mathbb{R}^{n\times n}$ such that $L=U\Lambda U^T$ where $\Lambda=\diag(\lambda_l)$. We can now define the graph Fourier transform (GFT) and its inverse \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}. The GFT of a spatial signal $x\in\mathbb{R}^n$ is $\hat{x}=U^Tx\in\mathbb{R}^n$, and the inverse GFT is $x=U\hat{x}$. Using GFT we can now generate fundamental operations of signal processing on graphs such as convolution and filtering. \\
{\bf Convolutional Filters.} The graph convolutional operator is simply defined in the Fourier space such as $f\ast_G g=U((U^Tf) \odot (U^Tg))$, where $\odot$ the element-wise Hadamard product. Then the convolutional filtering operation on graphs is defined as $y=g_\theta(L)x\in\mathbb{R}^n$, where $g_\theta(L)=g_\theta(U\Lambda U^T)=Ug_\theta(\Lambda) U^T\in\mathbb{R}^{n\times n}$ where $g_\theta(\Lambda)=\diag(\theta)$ and $\theta\in\mathbb{R}^n$ is the filter expressed in the Fourier space. Remember that convolutional filters, i.e. translation-invariant filters on graphs, can only be defined in the Fourier domain as convolution operators are not defined in the vertex space. \\

>>>>

%\usepackage{nips_2016}
\usepackage[final]{nips2016}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[english]{babel}
\usepackage[colorlinks=true,allcolors=blue]{hyperref}       % hyperlinks
%\usepackage{url}            % simple URL typesetting
%\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
%\usepackage{amssymb}
\usepackage{xfrac}       % compact symbols for 1/2, etc.
%\usepackage{microtype}      % microtypography
\usepackage{subfigure}
\usepackage{graphicx}
%\usepackage{tabularx}
\usepackage{booktabs}
%\usepackage{multirow}
%\usepackage{svg}
\usepackage{setspace}
\usepackage{color}

% Use less space
%\usepackage{titlesec}		% space before / after headings
%\titlespacing\section{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
%\titlespacing\subsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
%\titlespacing\subsubsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
%\usepackage{parskip} \setlength{\parskip}{5pt}
\setlength{\bibsep}{0.2pt}

@@

\author{
  Michaël Defferrard \\ %\thanks{further info} \\
  %Department of Electrical Engineering \\
  %École Polytechnique Fédérale de Lausanne (EPFL) \\
  EPFL, Lausanne, Switzerland \\
  \texttt{michael.defferrard@epfl.ch} \\
  \And % \AND
  Xavier Bresson \\
  EPFL, Lausanne, Switzerland \\
  \texttt{xavier.bresson@epfl.ch} \\
  \And % \AND
  Pierre Vandergheynst \\
  EPFL, Lausanne, Switzerland \\
  \texttt{pierre.vandergheynst@epfl.ch} \\
}

\usepackage[acronym]{glossaries}
\newacronym{CNN}{CNN}{Convolutional Neural Network}
\newacronym{SVD}{SVD}{Singular Value Decomposition}
\newacronym{SGD}{SGD}{Stochastic Gradient Descent}
\newacronym{PSD}{PSD}{positive semidefinite}

\DeclareMathOperator*{\diag}{diag}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\spn}{span}
\newcommand{\G}{\mathcal{G}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\bO}{\mathcal{O}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\tabref}[1]{Table~\ref{tab:#1}}
\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\eqnref}[1]{(\ref{eq:#1})}
\newcommand{\bruna}{art:BrunaZarembaSzlamLeCun13DLgraphs,
art:HenaffBrunaLeCun15DLgraphs}

\newcommand{\todo}[1]{{\color{red} #1 }}

\begin{document}

@@

\begin{abstract}

	Convolutional neural networks (CNNs) have greatly improved state-of-the-art
	performances in a number of fields, notably computer vision and natural
	language processing. In this work, we are interested in generalizing the
	formulation of CNNs from low-dimensional regular Euclidean domains, where
	image (2D), video (3D) and audio (1D) are represented, to high-dimensional
	irregular domains represented by graphs. This paper introduces a formulation
	of CNNs on graphs in the context of spectral graph theory. We borrow the
	fundamental tools from the emerging field of signal processing on graphs,
	which provides the necessary mathematical background and efficient numerical
	schemes to design localized graph filters which are fast to evaluate.
	Similarly to classical CNNs, the computational complexity as well as the
	number of parameters to learn are independent of the data dimensionality.
	Numerical experiments on MNIST and 20NEWS demonstrate the ability of the
	model to learn local and stationary features on graphs, as long as the graph
	is well-constructed. This work can be thought as the foundation of a second
	generation of CNNs on graphs built on spectral graph theory, being
	mathematically well defined, bringing better test performance and being
	faster to evaluate and train than the first generation introduced by Bruna,
	LeCun et al.

	%there is nothing to lose by going to the graph domain
	%except the rotation invariance

	%Graphs can either be the natural domain of the data, such as social networks,
	%biological graphs like gene regulatory and brain connectivity networks,
	%telecommunication networks, or constructed from the data, such as similarity
	%graphs.

	%not a generalization because it provides more invariance, but we don't want
	%to say it.
	%Our formulation is not a direct generalization of classical CNNs, in the
	%sense that it is not a spatial construction.
\end{abstract}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{images/illustrationCNNgraphs.eps}
%\includegraphics[width=13.5cm]{images/illustrationCNNgraphs.png}
\caption{Architecture of the proposed CNNs on graphs. Notation: $l$ is the
coarsening level, $x^l$ are the downsampled signals at layer $l$, $\G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the coarsening exponent, $n_c$ is the number of classes, $y$ is the output signal, and $\theta^l$ is the number of parameters to learn at $l$.}
\label{fig:}
\end{figure}

\section{Introduction}

Convolutional neural networks (CNNs) \cite{pro:LeCunBottouBengioHaffner98MNIST}
offer an efficient architecture to extract highly meaningful statistical
patterns in large-scale and high-dimensional datasets. The key success of CNNs
is its advanced ability to learn local stationary structures and compose them to
form multi-scale hierarchical patterns. Precisely, CNNs extract the local
stationarity property of the input data or signals by revealing local features
that are shared across the data domain. These similar features are identified
with localized convolutional filters or kernels, which are learned from the
data. Convolutional filters are shift- or translation-invariant filters, meaning
they are able to recognize identical features independently of their spatial
locations. Localized kernels or compactly supported filters refer to filters
that extract local features independently of the input data size, with a support
size that can be much smaller than the input size.

The compositional local stationarity property of CNNs can be efficiently
implemented on regular grids, like image and sound domains, as convolutional and
downsampling operators are mathematically well-defined on such discretized
Euclidean spaces. This has led to the breakthrough of CNNs in image, video, and
sound recognition tasks \cite{art:LeCunBengioHinton15DL} as these data lie on
low-dimensional regular lattices. But not all data lie on regular grids. User
data on social networks, gene data on biological regulatory networks, log data
on telecommunication networks, or text documents on words' embedding are
important examples of data lying on irregular or non-Euclidean domains.
Unfortunately, CNNs cannot be directly applied to such complex domains.

A generalization to irregular grids is not direct as the building blocks of
CNNs, the convolution and downsampling operators, are not mathematically clear
on these domains. This makes this extension challenging, both theoretically and
implementation-wise. This work leverages the recent works of
\cite{pro:GregorLeCun10LRF, pro:CoatesNg11LRF,
art:BrunaZarembaSzlamLeCun13DLgraphs, art:HenaffBrunaLeCun15DLgraphs,
art:HammondVandergheynstGribonval11GraphWav,
art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} to introduce an
efficient implementation of CNNs on irregular domains represented here as
graphs. Graphs are universal representations of heterogeneous pairwise
relationships between possibly high-dimensional data. Graphs can encode complex
geometric structures, and can be studied with strong mathematical tools such as
spectral graph theory, a blend between graph theory and harmonic analysis.

The major bottleneck of generalizing CNNs to graphs, and the primary goal of
this work, is the definition of localized filters which are efficient to
evaluate and learn.
%that are the core elements to extract local stationarity structures.
We will make use of recent tools developed in the context of graph signal
processing (GSP) \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}
to precisely control the spatial support of graph filters as well as to allow
fast evaluations. Eventually, a generic and efficient implementation of CNNs for
graph signals may pave the way to the development of a potentially large class
of deep learning techniques for graph-based data.

% Why convolution? Weight sharing, subject to assumptions
%\paragraph{Low learning complexity.} Learning deep fully connected networks is
%almost computationally intractable for large data domain.  Fortunately, for
%data satisfying stationarity and hierarchical compositionality structure, CNNs
%have proven to be remarkably powerful to identify highly informative features
%using low learning complexity. For example, learning a single layer with $n$
%input coordinates and $m<n$ outputs require $\bO(n^2)$ complexity to learn fully
%connected networks. It also requires the complexity $\bO(n)$ to learn any
%arbitrary filter, and $\bO(S)$ for localized filters, where $S$ is the size of
%the filter support, independent of the input size $n$. One of our main
%contributions is to stay with the same low complexity for learning graph
%convolutional filters. We will show it is actually possible to keep exactly the
%same complexity as standard CNNs, i.e. $\bO(S)$ for localized graph filters.

\todo{spatial approaches provide localization by definition}
\todo{why spectral approach?}
While a spatial approach is conceivable, it faces the challenge of matching
local neighborhoods, as shown by \cite{art:BrunaZarembaSzlamLeCun13DLgraphs}.

\todo{Another challenge is the definition of a pooling layer.}
Local stationarity property of data is extracted via localized convolutional
kernels. In this section, we are interested to extract the multi-scale
hierarchical composition property of data. This is efficiently achieved via grid
subsampling or downsampling and pooling, which trades spatial resolution with
feature resolution reducing the learning complexity without compromising the
system performances. In the case of standard CNNs for data lying on regular
grids, the subsampling operation is mathematically sound. The translation
operator commutes with the downsampling operator, hence not affecting the
translation invariance property of the system at different scales of grid
resolutions.

\paragraph{Graph Signal Processing.} GSP is an emerging field that aims at
bridging the gap between signal processing techniques like wavelet analysis and
graph theory such as spectral graph analysis
\cite{art:BelkinNiyogi05LaplaBeltrami, art:VonLuxburg07Tutorial}. A goal is to
generalize fundamental analysis operations for signals from regular grids to
irregular structures embodied by graphs. We refer the reader to
\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}
\todo{others for politics} for an
introduction of the field. Standard operations on grids such as convolution,
translation, filtering, dilatation, modulation or downsampling do not extend
directly to graphs and thus require new mathematical definitions while keeping
the original intuitive concepts. In this context, the authors of
\cite{art:HammondVandergheynstGribonval11GraphWav, art:CoifmanLafon06DifMap,
pro:GavishNadlerCoifman10GraphHaar} revisited the construction of wavelet
operators on graphs. In \cite{art:ShumanFarajiVandergheynst16PyramTrans,
art:RamEladCohen11TreeWavelets}, the authors designed a technique to perform
mutli-scale pyramid transforms on graphs. The works of
\cite{pro:TsitsveroBarbarossa15Uncert, pro:PasdeloupAlamiGriponRabbat15Uncert,
art:PerraudinRicaudShumanVandergheynst16Uncert} redefined uncertainty principles
on graphs, and showed that intuitive concepts may be lost, but can also produce
enhanced localization principles for signals on graphs. In
\cite{pro:HammondRaoaroorJacquesVandergheynst10LassoGraWav}, it was shown how to
carry out lasso-based signal regularization on graphs, and studied the
intertwined relationships between smoothness and sparsity on graphs. In
\cite{pro:TremblayPuyGribonvalVandergheynst16CompSpecClus}, the authors
investigated compressed sensing recovery conditions for graph spectral signal
analysis.

\todo{Model statistical assumptions: locality, stationarity, compositionality.}
These assumptions can be experimentally verified for any dataset. While we know
for them to be true for audio, images and videos, we show experimentally that
they might be satisfied for text documents as long as the graph is properly
constructed.
We also observe that
local stationarity is an assumption that is made about the data, and the CNNs
are fundamentally designed on this data hypothesis. However, we believe this
stationarity assumption to be present in many real-world data, as structured
data are usually composed of repetitive small patterns at different scales of
observation.

\section{Related Work}

%Extending the success of CNNs to
%non-Euclidean domains is essential to boost the power analysis of data lying on
%complex networks like biological, social or telecommunication networks.

Extending CNNs to non-Euclidean domains is a recent line of work. A first
attempt is the so-called local reception fields \cite{pro:GregorLeCun10LRF,
pro:CoatesNg11LRF}, successfully applied to image recognition. The main idea is
to group together features based upon a measure of similarity such as to select
a limited number of connections between two successive layers. While this model
reduces the number of parameters by exploiting the locality assumption, it does
not attempt to exploit any weight-sharing strategy.
%stationarity This approach is however not able to extract similar groups over
%the data domain, as it is not clear how to define convolutional filters on
%these groups.
\todo{read papers}

ShapeNet \cite{pro:MasciBoscainiBronsteinVandergheynst15GeoDL,
art:MasciBoscainiBronsteinVandergheynst15ShapeNet} is a generalization of CNNs
to 3D-meshes, a class of smooth low-dimensional non-Euclidean spaces. The
authors used geodesic polar coordinates to define mesh patches and formulated a
deep learning architecture which allows comparison across different manifolds.
They obtained state-of-the-art results for shape recognition tasks.

In \cite{pro:ChenChengMallat14deepHaar} and \cite{pro:RustamovGuibas14deepHaar},
the authors investigated the construction of Haar wavelet transforms on graphs
using a deep hierarchical architecture. They applied the method to object
recognition on sphere, and to sparse reconstruction of faces.
\todo{Learn parameters of a wavelet kernel? Then it is a parametric method,
they impose a shape to the filters.}
\todo{Unknown graph geometry?}

Finally, \cite{art:BrunaZarembaSzlamLeCun13DLgraphs} and
\cite{art:HenaffBrunaLeCun15DLgraphs} introduced a generalization of CNNs on
graphs using graph counterparts of grid convolution and downsampling. They were
able to learn convolutional filters on graphs using the convolution theorem
\cite{book:Mallat99wavelets} that defines convolutions as linear operators that
diagonalize in the Fourier basis (represented by the eigenvectors of the
Laplacian operator). They were thus the first to introduce a spectral
formulation to extend CNNs on graphs. They also introduced a strategy to learn
the graph structure from the data, and applied their model on image recognition,
text categorization and bioinformatics. This approach does however not scale due
to the necessary multiplications by the graph Fourier basis, a matrix of $n^2$
coefficients, where $n$ is the data dimensionality. Despite the cost of
computing this matrix, which requires an EVD on the graph Laplacian, the
dominant cost is the need to multiply the data by this matrix twice (forward and
inverse Fourier transforms) at a cost of $\bO(n^2)$ operations per forward and
backward pass, a computational bottleneck already identified by the authors.
Besides, as they rely on smoothness (via a spline parametrization) in the
Fourier domain to bring localization in the vertex domain, their model does not
provide a precise control over the local support of their kernels, which is
essential to learn localized filters. This work directly builds on theirs, and
we will show how we overcame these limitations and what we further gained.

\subsection{Contributions}

Our contributions can be summed up as follows:
\begin{enumerate}
%\setlength{\itemsep}{0pt} \setlength{\parskip}{0pt}
\item \textbf{Spectral formulation.} A clean spectral formulation built on
	established tools in GSP.
\item \textbf{Strictly localized filters.} Unlike \cite{\bruna}, the proposed spectral
	filters are strictly localized in a ball of radius $K$, i.e. $K$ hops from
	the central vertex.
\item \textbf{Low computational complexity.} The evaluation of our filters is
	linear w.r.t. the filters support's size $K$ and the number of edges $n \leq
	|\E| \leq n^2$. As most real-world graphs are highly sparse, we have $|\E|
	\ll n^2$ and $|\E| = kn$ for the widespread $k$-neighbors graphs. Moreover,
	this method avoids the Fourier basis altogether, thus the expensive
	eigenvalue decomposition (EVD) necessary to compute it as well as the need
	to store the basis, a matrix of size $n^2$. That is especially relevant when
	working with limited GPU memory. Besides the data, our method only requires
	to store the Laplacian, a sparse matrix of $|\E|$ non-zero values.
\item \textbf{Efficient pooling.} We propose an efficient pooling strategy on
	graphs which, after a rearrangement of the vertices, is analog to pooling of
	1D signals.
\item \textbf{Experimental results.} We present many experiments that ultimately
	show that our formulation is (1) a useful model, (2) computationally
	efficient and (3) superior both in accuracy and complexity to the previous
	generation of spectral graph CNNs introduced in \cite{\bruna}. We also show that
	our formulation performs similarly as a classical CNN on MNIST and study the
	impact of various graph constructions on performance.
\end{enumerate}

Beyond this work, our long-term goal is to make use of GSP tools as much as
possible to formulate mathematically well-designed techniques for CNNs on
graphs. This paper introduces the mathematical and computational foundations of
such models, while future developments may consider to enhance the building
blocks, namely graph filtering and coarsening. In the future, we expect to
further benefit from newly developed tools in GSP to enhance our formulation of
CNNs on graphs.

\section{Learning Fast Localized Spectral Filters} \label{sec:filters}

As there is no notion of spatial translation on graphs, convolutions are
implemented as multiplications in the spectral domain
\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}. While a spatial
convolution on an Euclidean domain is localized by definition (via the finite
size of the kernel) and has a linear computational complexity $\bO(n)$, a filter
defined in the spectral domain is not localized by default and costs $\bO(n^2)$
operations because of the necessary multiplication by the graph Fourier basis.
Note that while classical CNNs can be speed-up by leveraging the Fourier basis
(instead of computing the convolutions in the spatial domain)
\cite{mathieu_fast_2013}, it is only advantageous because of the $\bO(n \log(n))$
complexity of the FFT, which has no equivalent on graphs (yet). A good choice of
filter parametrization however allows us to prove that when translated (via a
convolution with a delta) to a given center vertex, the filter is strictly
localized in a ball of radius $K$ around that center vertex. Moreover, a
recursive formulation allows the filtering operation to scale linearly with $n$.
